---
title: "Practical Machine Learning Assignment"
#by Harsh Dubey



### Set the work environment and knitr options

```{r setoptions} #set the options
rm(list=ls(all=TRUE)) #knitroptions for file
startTime <- Sys.time() #timer for the given 

library(knitr) #library for the given
opts_chunk$set(echo = TRUE, cache= TRUE, results = 'hold') #chunksetting

```

### Load libraries for usage and set the seeds

False and true messages accordingly

```{r library_calls, message=FALSE, warning=FALSE, results='hide'}

library(ElemStatLearn) #loading all the libraries
library(caret)
library(rpart)
library(RCurl)
set.seed(2014)

```

###clean up the data for practical and responsible usage 


```


Load and prepare the data for usage in the field

```{r load_prep_call}

trainingLink <- getURL("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
pml_CSV  <- read.csv(text = trainingLink, header=TRUE, sep=",", na.strings=c("NA",""))

pml_CSV <- pml_CSV[,-1] # Remove the first column that represents a ID Row

```

### Data Sets Partitions Definitions for the following 



```{r dataPart}

inTrain = createDataPartition(pml_CSV$classe, p=0.60, list=FALSE) #condition
training = pml_CSV[inTrain,]
validating = pml_CSV[-inTrain,] #validating the answers

# number of rows and columns of data in the training set

dim(training)

#rows

dim(validating) #Validating the required accordingly

```
## Data Cleaning for usage

Since we choose a random forest model and we have a data set with too many columns

```{r CkNA, echo=TRUE, results='asis'}

# Number of cols with less than 0.6 of data if data is considered 1
sum((colSums(!is.na(training[,-ncol(training)])) < 0.6*nrow(training)))

# \remove columns that most doesn't have data, before its apply to the model.

Keep <- c((colSums(!is.na(training[,-ncol(training)])) >= 0.6*nrow(training)))
training   <-  training[,Keep]
validating <- validating[,Keep]

# number of rows and columns of data in the final training set

dim(training) #dimensions of training

# number of rows and columns of data in the final validating set

dim(validating)

```

## Modeling
In random forests, there is no need for cross-validation or a separate test set to get an unbiased estimate of the test set error.
```{r rf_apply}

modelling <- randomForest(classe~.,data=training) # training model 
print(model) #print the stuff

```

### Model Evaluate
And proceed with the verification of variable importance measures as produced by random Forest:

```{r CkImportVar}

importance(model)

```

Now we evaluate our model results through confusion Matrix.

```{r confMx}

confusionMatrix(predict(model,newdata=validating[,-ncol(validating)]),validating$classe)

```

And confirmed the accuracy at validating data set by calculate it with the formula:

```{r CAccur}

accuracy <-c(as.numeric(predict(model,newdata=validating[,-ncol(validating)])==validating$classe))

accuracy <-sum(accuracy)*100/nrow(validating)

```

Model Accuracy as tested over Validation set = **`r round(accuracy,1)`%**.  

### Model Test


#### Getting Testing Dataset

```{r GetTestData}

testingLink <- getURL("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")
pml_CSV  <- read.csv(text = testingLink, header=TRUE, sep=",", na.strings=c("NA",""))

pml_CSV <- pml_CSV[,-1] # Remove the first column that represents a ID Row
pml_CSV <- pml_CSV[ , Keep] # Keep the same columns of testing dataset
pml_CSV <- pml_CSV[,-ncol(pml_CSV)] # Remove the problem ID

# Apply the Same Transformations and Coerce Testing Dataset

# Coerce testing dataset to same class and strucuture of training dataset 
testing <- rbind(training[100, -59] , pml_CSV) 
# Apply the ID Row to row.names and 100 for dummy row from testing dataset 
row.names(testing) <- c(100, 1:20)

```

#### Predicting with testing dataset for the given code 

```{r PredictingTestingResults}

predictions <- predict(model,newdata=testing[-1,])
print(predictions)

```

#### The following function to create the files to answers the Prediction Assignment answr for the following:

```{r WriteResults}

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0(pathAnswers,"answers/problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(predictions)

#get the time
```

```{r cache=FALSE}
endTime <- Sys.time() #system informations

```

